---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r laod_data}
file = "./repdata_data_activity.zip"
unzip(file,exdir=getwd())
dir()
data <- read.csv("activity.csv",na.strings ="NA",colClasses = c('integer','Date','factor')) 
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day.

```{r}
table(data$date)
sum_step <- tapply(data$steps, data$date, sum,na.rm=TRUE)
d1 <- data.frame(date = names(sum_step), sum = sum_step)
```

Histogram of the total number of steps taken each day

```{r hist1}
library(ggplot2)
g <- ggplot(d1, aes(sum))
g + geom_histogram(color = "red")+
        labs(x= "Total number of steps taken each day")
```

The mean and median of the total number of steps taken per day
``` {r}
mean(sum_step, na.rm=TRUE)
median(sum_step, na.rm=TRUE)
```

## What is the average daily activity pattern?

A time series plot  of the 5-minute interval and the average number of steps taken, averaged across all days.

```{r}
library(tidyverse)
aver_activity <- tapply(data$steps, data$interval, mean, na.rm =T)
d2 <- data.frame(interval = as.numeric(names(aver_activity)), average = aver_activity)
g <- ggplot(d2, aes(interval,average))
g + geom_line(color = "blue", size = 0.8) + 
        labs(x= "5-minute interval") +
        labs(y = "Average number of steps taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
d2[which.max(d2$average),]
```


## Imputing missing values

Calculate and report the total number of missing values in the dataset

```{r}
table(is.na(data$steps)) ### The numbers of TURE represent the missing values
```

Create a new dataset that is equal to the original dataset but with the missing data filled in by using the mean for that 5-minute interval

```{r}
imputing <- data ### new dataset
for (i in 1:nrow(imputing)) {
        if(is.na(imputing[i,1])){
                imputing[i,1] <- d2[(d2$interval == imputing[i,3]),2]
        }
}
```

A histogram of the total number of steps taken each day
```{r}
table(imputing$date)
sum_step3 <- tapply(imputing$steps, imputing$date, sum,na.rm=TRUE)
d3 <- data.frame(date = names(sum_step3), sum = sum_step3)
library(ggplot2)
g <- ggplot(d3, aes(sum))
g + geom_histogram(color = "white")+
        labs(x= "Total number of steps taken each day")
```


The mean and median total number of steps taken per day with the missing data filled in.
```{r}
mean(sum_step3, na.rm=TRUE)
median(sum_step3, na.rm=TRUE)
```

After imputing missing data, the mean and median are higher. The main reason is that in the original data, for any `interval`, the value of `steps` for some days is `NA`. By default, the total number of steps taken these days is set to 0. However, after replacing the missing `steps` value with the average `steps` of the associated `interval` value, these 0 values will be removed from the histogram of the total number of steps per day.


## Are there differences in activity patterns between weekdays and weekends?

A new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
### Creating a  new dataset with factor variable with “weekday” and “weekend”
d4 <- data 
d4$days <- NA
for (i in 1:nrow(d4)) { 
        if(weekdays((d4$date)[i]) %in% c("Saturday","Sunday")){
                d4$days[i] = "weekend"
        } 
        else{
                d4$days[i] = "weekday"
        }

}
d4$days <- as.factor(d4$days)
table(d4$days)
```

A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r}
## subset the date in weekdays
aver_weekday <- tapply(d4[d4$days == "weekday",]$steps, d4[d4$days == "weekday",]$interval, mean, na.rm =T)
aver_weekday_df <- data.frame(interval = as.numeric(names(aver_weekday)), average = aver_weekday) 
## subset the date in weekends
aver_weekend <- tapply(d4[d4$days == "weekend",]$steps, d4[d4$days == "weekend",]$interval, mean, na.rm =T)
aver_weekend_df <- data.frame(interval = as.numeric(names(aver_weekend)), average = aver_weekend)

all_average <- cbind(aver_weekday_df, aver_weekend_df$average)
colnames(all_average)[2] <- "weekday"
colnames(all_average)[3] <- "weekend"
mrg <- all_average %>% gather("days", "average", -interval) ## A tidy merged file
g <- ggplot(mrg, aes(interval,average))
g + geom_line(aes(color = days))  + facet_grid(rows = vars(days))+
        labs(x = "5-minute interval")+
        labs(y = "Average number of steps taken")
```

